\input{preamble}

% OK, start here.
%
\begin{document}

\title{Algebraic Spaces}


\maketitle

\phantomsection
\label{section-phantom}

\tableofcontents

\section{Introduction}
\label{section-introduction}

\noindent
This is where we define algebraic spaces and make some very elementary
observations. Algebraic spaces were first introduced by Mike Artin,
see \cite{ArtinI} and \cite{ArtinII}.





\section{General remarks}
\label{section-general}

\noindent
We work in a suitable big fppf site $\textit{Sch}_{fppf}$
as in Topologies, Definition \ref{topologies-definition-big-fppf-site}.
So, if not explicitly stated otherwise all schemes will be objects
of $\textit{Sch}_{fppf}$.
We will record elsewhere what changes if you change the big
fppf site (insert future reference here).

\medskip\noindent
We will always work relative to a base $S$ contained in $\textit{Sch}_{fppf}$.
And we will then work with the big fppf site $(\textit{Sch}/S)_{fppf}$,
see Topologies, Definition \ref{topologies-definition-big-small-fppf}.
The absolute case can be recovered by taking
$S = \text{Spec}(\mathbf{Z})$.

\medskip\noindent
If $U, T$ are schemes over $S$, then we denote
$U(T)$ for the set of $T$-valued points {\it over} $S$.
In a formula: $U(T) = \text{Mor}_S(T, U)$.

\medskip\noindent
Note that any fpqc covering is a family of universally effective
epimorphisms, see
Descent, Lemma \ref{fpqc-descent-lemma-fpqc-universal-effective-epimorphisms}.
Hence the topology on $\textit{Sch}_{fppf}$
is weaker than the canonical topology and all representable presheaves
are sheaves.







\section{Representable morphisms of presheaves}
\label{section-representable}

\noindent
Let $S$ be a scheme contained in $\textit{Sch}_{fppf}$.
Let $F, G : (\textit{Sch}/S)_{fppf}^{opp} \to \textit{Sets}$.
Let $a : F \to G$ be a representable transformation of functors, see
Categories,
Definition \ref{categories-definition-representable-map-presheaves}.
This means that for every
$U \in \text{Ob}((\textit{Sch}/S)_{fppf})$ and
any $\xi \in G(U)$ the fiber product $h_U \times_{\xi, G} F$ is representable.
Choose a representing object $V_\xi$ and an isomorphism
$h_{V_\xi} \to h_U \times_G F$.
By the Yoneda lemma, see Categories, Lemma \ref{categories-lemma-yoneda},
the projection $h_{V_\xi} \to h_U \times_G F \to h_U$ comes from a unique
morphism of schemes $a_\xi : V_\xi \to U$.
Suggestively we could represent this by the diagram
$$
\xymatrix{
V_\xi \ar@{~>}[r] \ar[d]_{a_\xi} & h_{V_\xi} \ar[d] \ar[r] & F \ar[d]^a \\
U \ar@{~>}[r] & h_U \ar[r]^\xi & G
}
$$
where the squiggly arrows represent the Yoneda embedding.

\begin{definition}
\label{definition-relative-representable-property}
With $S$, and $a : F \to G$ representable as above.
Let $P$ be a property of morphisms of schemes which
\begin{enumerate}
\item is preserved under any base change,
see Schemes, Definition \ref{schemes-definition-preserved-by-base-change},
and
\item is fppf local on the base, see
Descent, Definition \ref{fpqc-descent-definition-property-local}.
\end{enumerate}
In this case we say that $a$ has {\it property $P$} if for every
$U \in \text{Ob}((\textit{Sch}/S)_{fppf})$ and
any $\xi \in G(U)$ the resulting morphism of schemes
$V_\xi \to U$ has property $P$.
\end{definition}

\noindent
It is important to note that we will only use this definition for
properties of morphisms that are stable under base change. This is
not because the definition doens't make sense otherwise; rather it
is because we may want to give a different definition which is
better suited to the property we have in mind.

\begin{remark}
\label{remark-list-properties-stable-base-change}
Here is a list of properties/types of morphisms
which are stable under arbitrary base change:
\begin{enumerate}
\item (closed, or open) immersions, see
Schemes, Lemma \ref{schemes-lemma-base-change-immersion},
\item quasi-compact, see
Schemes, Lemma \ref{schemes-lemma-quasi-compact-preserved-base-change},
\item universally closed, see
Schemes, Definition \ref{schemes-definition-universally-closed},
\item (quasi-)separated, see
Schemes, Lemma \ref{schemes-lemma-separated-permanence},
\item surjective, see
Morphisms, Lemma \ref{morphisms-lemma-base-change-surjective},
\item radicial (or universally injective), see
Morphisms, Lemma \ref{morphisms-lemma-radicial-universally-injective},
\item affine, see
Morphisms, Lemma \ref{morphisms-lemma-base-change-affine},
\item quasi-affine, see
Morphisms, Lemma \ref{morphisms-lemma-base-change-quasi-affine},
\item (locally) of finite type, see
Morphisms, Lemma \ref{morphisms-lemma-base-change-finite-type},
\item (locally) quasi-finite, see
Morphisms, Lemma \ref{morphisms-lemma-base-change-quasi-finite},
\item (locally) of finite presentation, see
Morphisms, Lemma \ref{morphisms-lemma-base-change-finite-presentation},
\item locally of finite type of relative dimension $d$, see
Morphisms, Lemma \ref{morphisms-lemma-base-change-relative-dimension-d},
\item universally open, see
Morphisms, Definition \ref{morphisms-definition-open},
\item flat, see
Morphisms, Lemma \ref{morphisms-lemma-base-change-flat},
\item syntomic, see
Morphisms, Lemma \ref{morphisms-lemma-base-change-syntomic},
\item smooth, see
Morphisms, Lemma \ref{morphisms-lemma-base-change-smooth},
\item unramified, see
Morphisms, Lemma \ref{morphisms-lemma-base-change-unramified},
\item etale, see
Morphisms, Lemma \ref{morphisms-lemma-base-change-etale},
\item proper, see
Morphisms, Lemma \ref{morphisms-lemma-base-change-proper},
\item H-projective, see
Morphisms, Lemma \ref{morphisms-lemma-H-projective-base-change},
\item (locally) projective, see
Morphisms, Lemma \ref{morphisms-lemma-base-change-projective},
\item finite or integral, see
Morphisms, Lemma \ref{morphisms-lemma-base-change-finite},
\item finite locally free, see
Morphisms, Lemma \ref{morphisms-lemma-base-change-finite-locally-free}.
\end{enumerate}
Add more as needed.
\end{remark}

\begin{remark}
\label{remark-list-properties-relative-representable-transformations}
Of the properties mentioned in Remark
\ref{remark-list-properties-stable-base-change}
the following are also fpqc local on the base
(and a fortiori fppf local on the base):
\begin{enumerate}
\item for immersions we have this for
\begin{enumerate}
\item closed immersions, see
Descent, Lemma \ref{fpqc-descent-lemma-descending-property-closed-immersion},
\item open immersions see
Descent, Lemma \ref{fpqc-descent-lemma-descending-property-open-immersion}, and
\item quasi-compact immersions, see
Descent,
Lemma \ref{fpqc-descent-lemma-descending-property-quasi-compact-immersion},
\end{enumerate}
\item quasi-compact, see
Descent, Lemma \ref{fpqc-descent-lemma-descending-property-quasi-compact},
\item universally closed, see
Descent,
Definition \ref{fpqc-descent-lemma-descending-property-universally-closed},
\item (quasi-)separated, see
Descent, Lemmas
\ref{fpqc-descent-lemma-descending-property-quasi-separated}, and
\ref{fpqc-descent-lemma-descending-property-separated},
\item surjective, see
Descent, Lemma \ref{fpqc-descent-lemma-descending-property-surjective},
\item radicial (or universally injective), see
Descent, Lemma \ref{fpqc-descent-lemma-descending-property-radicial},
\item affine, see
Descent, Lemma \ref{fpqc-descent-lemma-descending-property-affine},
\item quasi-affine, see
Descent, Lemma \ref{fpqc-descent-lemma-descending-property-quasi-affine},
\item (locally) of finite type, see
Descent,
Lemmas \ref{fpqc-descent-lemma-descending-property-locally-finite-type}, and
\ref{fpqc-descent-lemma-descending-property-finite-type},
\item (locally) quasi-finite, see
Descent, Lemma \ref{fpqc-descent-lemma-descending-property-quasi-finite},
\item (locally) of finite presentation, see
Descent, Lemmas
\ref{fpqc-descent-lemma-descending-property-locally-finite-presentation}, and
\ref{fpqc-descent-lemma-descending-property-finite-presentation},
\item locally of finite type of relative dimension $d$, see
Descent,
Lemma \ref{fpqc-descent-lemma-descending-property-relative-dimension-d},
\item universally open, see
Descent, Lemma \ref{fpqc-descent-lemma-descending-property-universally-open},
\item flat, see
Descent, Lemma \ref{fpqc-descent-lemma-descending-property-flat},
\item syntomic, see
Descent, Lemma \ref{fpqc-descent-lemma-descending-property-syntomic},
\item smooth, see
Descent, Lemma \ref{fpqc-descent-lemma-descending-property-smooth},
\item unramified, see
Descent, Lemma \ref{fpqc-descent-lemma-descending-property-unramified},
\item etale, see
Descent, Lemma \ref{fpqc-descent-lemma-descending-property-etale},
\item proper, see
Descent, Lemma \ref{fpqc-descent-lemma-descending-property-proper},
\item finite or integral, see
Descent, Lemma \ref{fpqc-descent-lemma-descending-property-finite},
\item finite locally free, see
Descent,
Lemma \ref{fpqc-descent-lemma-descending-property-finite-locally-free}.
\end{enumerate}
Add more here as needed.
\end{remark}

\begin{lemma}
\label{lemma-representable-diagonal}
Let $S$ be a scheme contained in $\textit{Sch}_{fppf}$.
Let $F$ be a presheaf of sets on $(\textit{Sch}/S)_{fppf}$.
The following are equivalent:
\begin{enumerate}
\item The diagonal $F \to F \times F$ is representable.
\item For every scheme $U$ over $S$,
$U/S \in \text{Ob}((\textit{Sch}/S)_{fppf})$
and any $\xi \in F(U)$ the map $\xi : h_U \to F$ is representable.
\end{enumerate}
\end{lemma}

\begin{proof}
This is completely formal, see
Categories, Lemma \ref{categories-lemma-representable-diagonal}.
It depends only on the fact that the category $(\textit{Sch}/S)_{fppf}$
has products of pairs of objects and fibre products, see
Topologies, Lemma \ref{topologies-lemma-fibre-products-fppf}.
\end{proof}

\noindent
In the situation of the lemma, for any morphism
$\xi : h_U \to F$ as in the lemma, it makes sense
to say that $\xi$ has property $P$, for any property
as in Definition \ref{definition-relative-representable-property}.
In particular this holds for $P = $ ``surjective'' and $P = $ ``etale'', see
Remark \ref{remark-list-properties-relative-representable-transformations}
above. We will use these in the definition
of algebraic spaces below. 


















\section{Algebraic spaces}
\label{section-algebraic-spaces}


\begin{definition}
\label{definition-algebraic-space}
Let $S$ be a scheme contained in $\textit{Sch}_{fppf}$.
An {\it algebraic space over $S$} is a presheaf
$$
F : (\textit{Sch}/S)^{opp}_{fppf} \longrightarrow \textit{Sets}
$$
with the following properties
\begin{enumerate}
\item The presheaf $F$ is a sheaf.
\item The diagonal morphism $F  \to F \times F$ is representable.
\item There exists a scheme $U \in \text{Ob}(\textit{Sch}_{fppf})$
and a map $h_U \to F$ which is surjective, and etale.
\end{enumerate}
\end{definition}

\noindent
There are two differences with the ``usual'' definition, for example the
definition in Knutson's book \cite{Kn}.

\medskip\noindent
The first is that we require $F$ to be a sheaf in the fppf topology.
One reason for doing this is that many natural examples
of algebraic spaces satisfy the sheaf condition for the fppf coverings
(and even for fpqc coverings). Also, one of the reasons that algebraic
spaces have been so useful is via Mike Artin's results on algebraic spaces.
Built into his method is a condition which garantees the result is
locally of finite presentation over $S$.
Combined it somehow seems to us that the fppf topology
is the natural topology to work with. In the end the resulting category
of algebraic spaces ends up being ``the same''. Namely, allthough the actual
sheaves $F$ being considered will actually be different in the end the
category of algebraic spaces defined using sheaves in the etale topology
is equivalent the the category we define here. This will be clear later
when we introduce presentations (insert future reference here).

\medskip\noindent
The second is that we only require the diagonal map for $F$ to be
representable, whereas in \cite{Kn} it is required that it also
be quasi-compact. If $F = h_U$ for some scheme $U$ over $S$
this corresponds to the condition that $S$ be quasi-separated.
Our point of view is to try to prove a certain
number of the results that follow only assuming that the diagonal
of $F$ be representable, and simply add an addition hypothesis wherever
this is necessary. In any case it has the pleasing consquence that
the following lemma is true.

\begin{lemma}
\label{lemma-scheme-is-space}
A scheme is an algebraic space. More precisely,
given a scheme $T \in \text{Ob}((\textit{Sch}/S)_{fppf})$
the representable functor $h_T$ is an algebraic space.
\end{lemma}

\begin{proof}
The functor $h_T$ is a sheaf by our remarks in Section \ref{section-general}.
The diagonal $h_T \to h_T \times h_T = h_{T \times T}$ is
representable because $(\textit{Sch}/S)_{fppf}$ has fibre products.
The identity map $h_T \to h_T$ is surjective etale.
\end{proof}

\begin{definition}
\label{definition-morphism-algebraic-spaces}
Let $F$, $F'$ be algebraic spaces over $S$.
A {\it morphism $f : F \to F'$ of algebraic spaces over $S$}
is a transformation of functors from $F$ to $F'$.
\end{definition}

\noindent
The category of algebraic spaces over $S$ contains the category
$(\textit{Sch}/S)_{fppf}$ as a full subcategory via the
Yoneda embedding $T/S \mapsto h_T$. From now on we no longer distinghuish
between a scheme $T/S$ and the algebraic space it represents.
Thus when we say ``Let $f : T \to F$ be a morphism from the scheme
$T$ to the algebraic space $F$'', we mean that
$T \in \text{Ob}((\textit{Sch}/S)_{fppf})$, that $F$ is an
algebraic space over $S$, and that $f : h_T \to F$ is a morphism
of algebraic spaces over $S$.








\section{Glueing algebraic spaces}
\label{section-glueing-algebraic-spaces}

\noindent
In this section we really start abusing notation and not
distinguish between schemes and the spaces they represent.

\begin{lemma}
\label{lemma-coproduct-representable-sheaf}
Let $S \in \text{Ob}(\textit{Sch}_{fppf})$.
Let $U \in \text{Ob}(\textit{Sch}/S)_{fppf}$.
Given a set $I$ and sheaves $F_i$ on $\text{Ob}(\textit{Sch}/S)_{fppf}$,
if $U \cong \coprod_{i\in I} F_i$
as sheaves, then each $F_i$ is representable by an open and closed
subscheme $U_i$ and $U \cong \coprod U_i$ as schemes.
\end{lemma}

\begin{proof}
By assumption this means there exists an fppf covering
$\{U_j \to U\}_{j \in J}$ such that each $U_j \to U$
factors through $F_{i(j)}$ for some $i(j) \in I$.
Denote $V_j = \text{Im}(U_j \to U)$.
This is an open of $U$ by
Morphisms, Lemma \ref{morphisms-lemma-fppf-open}, and
$\{U_j \to V_j\}$ is an fppf covering. Hence it follows that
$V_j \to U$ factors through $F_{i(j)}$ since $F_{i(j)}$ is
a subsheaf. It follows from $F_i \cap F_{i'} = \empty$, $i \not = i'$
that $V_j \cap V_{j'} = \emptyset$
unless $i(j) = i(j')$. Hence we can take
$U_i = \bigcup_{j,\ i(j) = i} V_j$ and everything is clear.
\end{proof}

\noindent
Here is the analogue of Schemes, Lemma \ref{schemes-lemma-glue-functors}.

\begin{lemma}
\label{lemma-glueing-algebraic-spaces}
Let $S \in \text{Ob}(\textit{Sch}_{fppf})$.
Let $F$ be a presheaf of sets on $(\textit{Sch}/S)_{fppf}$.
Assume
\begin{enumerate}
\item $F$ is a sheaf,
\item there exists an index set $I$
and subfunctors $F_i \subset F$ such that
\begin{enumerate}
\item $\coprod F_i$ is an algebraic space\footnote{
This basically just means each $F_i$ is an algebraic space, see
Lemma \ref{lemma-coproduct-representable-sheaf}.},
\item each $F_i \to F$ is a representable,
\item each $F_i \to F$ is an open immersion (see
Definition \ref{definition-relative-representable-property} and
Remark \ref{remark-list-properties-relative-representable-transformations}),
and
\item the map of sheaves $\coprod F_i \to F$ is surjective.
\end{enumerate}
\end{enumerate}
Then $F$ is an algebraic space.
\end{lemma}

\begin{proof}
Let $T$, $T'$ be objects of $(\textit{Sch}/S)_{fppf}$.
Let $T \to F$, $T' \to F$ morphisms.
The assumptions imply that there exists an open covering
$T = \bigcup V_i$ such that $V_i = T \times_F F_i$.
Note that this in particular implies that
$\coprod F_i \to F$ is surjective in the Zariski topology!
Also write similarly $T' = \bigcup V'_i$ with $V'_i = T' \times_F F_i$.

\medskip\noindent
To show that the diagonal $F \to F \times F$ is representable
we have to show that $G = T \times_F T'$ is representable.
Consider the subfunctors $G_i = G \times_F F_i$.
Note that $G_i = V_i \times_{F_i} V'_i$, and hence is representable
as $F_i$ is an algebraic space.
By the above the $G_i$ form a Zariski covering of $F$.
Hence by Schemes, Lemma \ref{schemes-lemma-glue-functors}
we see $G$ is representable.

\medskip\noindent
Choose a scheme $U \in \text{Ob}(\textit{Sch}/S)_{fppf}$
and a surjective
etale morphism $U \to \coprod F_i$ (this exist by hypothesis).
We may write $U = \coprod U_i$ with $U_i$ the inverse image of $F_i$,
see Lemma \ref{lemma-coproduct-representable-sheaf}.
We claim that $U \to F$ is surjective and
etale. Surjectivity follows as $\coprod F_i \to F$ is surjective.
Consider the fibre product $U \times_F T$ where $T \to F$ is as
above. We have to show that $U \times_F T \to T$ is etale.
Since $U \times_F T = \coprod U_i \times_F T$ it suffices to show
each $U_i \times_F T \to T$ is etale. Since
$U_i \times_F T = U_i \times_{F_i} V_i$ this follows from the
fact that $U_i \to F_i$ is etale and $V_i \to T$ is an open immersion
(and Morphisms, Lemmas \ref{morphisms-lemma-open-immersion-etale}
and \ref{morphisms-lemma-composition-etale}).
\end{proof}














\section{Presentations of algebraic spaces}
\label{section-presentations}

\noindent
Given an algebraic space we can find a ``presentation'' of it.


\begin{lemma}
\label{lemma-space-presentation}
Let $F$ be an algebraic space over $S$. Let $f : U \to F$ be a
surjective etale morphism from a scheme to $F$. Set $R = U \times_F U$.
Then
\begin{enumerate}
\item $j : R \to U \times_S U$ defines an equivalence relation on
$U$ over $S$ (see
Groupoids, Definition \ref{groupoids-definition-equivalence-relation}).
\item the morphisms $s, t : R \to U$ are etale, and
\item the diagram
$$
\xymatrix{
R \ar@<1ex>[r] \ar@<-1ex>[r] &
U \ar[r] &
F
}
$$
is a coequalizer diagram in $\textit{Sh}((\textit{Sch}/S)_{fppf})$.
\end{enumerate}
\end{lemma}

\begin{proof}
Let $T/S$ be an object of $(\textit{Sch}/S)_{fppf}$.
Then $R(T) = \{(a, b) \in U(T) \times U(T) \mid f \circ a = f \circ b\}$
which is clearly defines an equivalence relation on $U(T)$.
The morphisms $s, t : R \to U$ are etale because the morphism
$U \to F$ is etale.

\medskip\noindent
To prove (3) we first show that
$U \to F$ is a surjection of sheaves, see
Sites, Definition \ref{sites-definition-sheaves-injective-surjective}.
Let $\xi \in F(T)$ with $T$ as above. Let $V = T \times_{\xi, F, f}U$.
By assumption $V$ is a scheme and $V \to T$ is surjective etale.
Hence $\{V \to T\}$ is a covering for the fppf topology.
Since $\xi|_V$ factors through $U$ by construction we
conclude $U \to F$ is surjective. To conclude we
have to show that given any two morphisms
$a, b : T \to U$ such that $f \circ a = f \circ b$ there is a
morphism $c : T \to R$ such that $a = \text{pr}_0 \circ c$
and $b = \text{pr}_1 \circ b$. This is clear from the definition
of $R$.
\end{proof}

\noindent
This lemma suggests the following definitions.

\begin{definition}
\label{definition-etale-equivalence-relation}
Let $S$ be a scheme. Let $U$ be a scheme over $S$.
An {\it etale equivalence relation} on $U$ over $S$
is an equivalence relation $j : R \to U \times_S U$
such that $s, t : R \to U$ are etale morphisms of schemes.
\end{definition}

\begin{definition}
\label{definition-presentation}
Let $F$ be an algebraic space over $S$.
A {\it presentation} of $F$ is given by a scheme
$U$ over $S$ and an etale equivalence relation $R$ on $U$ over $S$, and
a surjective etale morphism $U \to F$ such that $R = U \times_F U$.
\end{definition}

\noindent
Equivalently we could ask for the existence of an isomorphism
$$
U/R \cong F
$$
where the quotient $U/R$ is as defined in
Groupoids, Section \ref{groupoids-section-representable-quotients}.

\medskip\noindent
To construct algebraic spaces we study the converse question, namely,
for which pairs $(U, R)$ the quotient sheaf $U/R$ is an algebraic
space.






























\section{Algebraic spaces and equivalence relations}
\label{section-spaces-from-equivalence-relations}

\noindent
Suppose given a scheme $U$ over $S$
and an equivalence relation $R$ on $U/S$ such that the two
projection morphisms $s, t : R \to U$ are etale.
We would like this to define an algebraic space.
Actually this may not always be the case.
The problem is that it is not clear that the sheafification
of the quotient sheaf $U/R$ has all the properties required
of it in Definition \ref{definition-algebraic-space}.

\begin{lemma}
\label{lemma-pullback-etale-equivalence-relation}
Let $S$ be a scheme. Let $U$ be a scheme over $S$.
Let $j = (s, t) : R \to U \times_S U$
be an etale equivalence relation on $U$ over $S$.
Let $U' \to U$ be an etale morphism.
Let $R'$ be the restriction of $R$ to $U'$, see
Groupoids, Definition \ref{groupoids-definition-restrict-relation}.
Then $j' : R' \to U' \times_S U'$ is an etale equivalence
relation also.
\end{lemma}

\begin{proof}
It is clear from the description of $s', t'$ in
Groupoids, Lemma \ref{groupoids-lemma-restrict-groupoid}
that $s' , t' : R' \to U'$ are etale
as compositions of base changes of etale morphisms
(see Morphisms, Lemma \ref{morphisms-lemma-base-change-etale}
and \ref{morphisms-lemma-composition-etale}).
\end{proof}

\begin{lemma}
\label{lemma-pre-presentation}
Let $S$ be a scheme.
Let $U$ be a scheme over $S$.
Let $j = (s, t) : R \to U \times_S U$ be an equivalence relation
on $U$ over $S$.
Let $g : U' \to U$ be a morphism, and let $R' = R|_{U'}$ be the
restriction of $R$ to $U$.
If $s, t : R \to U$ are universally open, and $g$ is flat and
locally of finite presentation, then $R'/U' \to R/U$ is
representable, and is an open immersion.
\end{lemma}

\begin{proof}
By Groupoids, Lemma \ref{groupoids-lemma-restrict-relation}
the morphism $j' = (t', s') : R' \to U' \times_S U'$
defines an equivalence relation.
Since $g$ is flat and locally of
finite presentation we see that $g$ is universally open as well
(Morphisms, Lemma \ref{morphisms-lemma-fppf-open}).
Hence, by the description of $R'$ in
Groupoids, Lemma \ref{groupoids-lemma-restrict-groupoid}
we see that $s', t'$ are universally open as well.
Let $W^1 = g(U') \subset U$, and let $W = t(s^{-1}(W^1))$.
Then $W^1$ and $W$ are open in $U$. Moreover, as $j$ is an
equivalence relation we have $t(s^{-1}(W)) = W$.

\medskip\noindent
By Groupoids, Lemma \ref{groupoids-lemma-quotient-groupoid-restrict}
the map of sheaves $F' = U'/R' \to F = U/R$ is injective.
Let $a : T \to F$ be a morphism from a scheme into $U/R$.
We have to show that $T \times_F F'$ is representable
by an open subscheme of $T$.

\medskip\noindent
The morphism $a$ is given by the following data:
an fppf covering $\{\varphi_j : T_j \to T\}_{j \in J}$ of $T$ and
morphsms $a_j : T_j \to U$ such that the maps
$$
a_j \times a_{j'} :
T_j \times_T T_{j'}
\longrightarrow
U \times_S U
$$
factor through $j : R \to U \times_S U$ via some (unique) maps
$r_{jj'} : T_j \times_T T_{j'} \to R$. The system
$(a_j)$ corresponds to $a$ in the sense that the diagrams
$$
\xymatrix{
T_j \ar[r]_{a_j} \ar[d] & U \ar[d] \\
T \ar[r]^a & U/R
}
$$
commute.

\medskip\noindent
Consider the open subsets $W_j = a_j^{-1}(W) \subset T_j$.
Since $t(s^{-1}(W)) = W$ we see that
$$
W_j \times_T T_{j'} =
r_{jj'}^{-1}(t^{-1}(W)) = r_{jj'}^{-1}(s^{-1}(W)) =
T_j \times_T W_{j'}.
$$
By Descent, Lemma \ref{} this means there exists a unique open
$W_T \subset T$ such that $\varphi_j^{-1}(W_T) = W_j$.
Since $T_j \to T$ is flat and of finite presentation the image $\varphi_j(W_j) \subset T$
is open (see Morphisms, Lemma \ref{morphisms-lemma-fppf-open}).
Set $W = \bigcup \varphi_j(W_j) \subset T$. Thinking of $W$ as
an open subscheme of $T$ we see that $a|_W$ factors through
$U'/R'$. Namely $\{\varphi_j|_{W_j} : W_j \to W\}$
is an fppf covering, each $a_j|_{W_j} : W_j \to U'$ and
of course $r_{jj'}|_{W_j \times_S W_{j'}}$ factors through $R'$.
We claim that for any morphism of schemes $f : T' \to T$
the element $a|_{T'} \in F(T')$ lies in $F'(T')$ if and
only if $f$ factors through $W$.

\medskip\noindent
We have seen the ``if'' part above.
On the other hand, if $a|_{T'} \in F'(T')$, then
there exists a refinement $\{T'_i \to T'\}_{i \in I}$ of
the covering $\{T' \times_T T_j \to T\}$, given by a map
$\alpha : I \to J$ and morphisms
$g_i : T'_i \to T' \times_T T_{\alpha(i)}$ over $T'$,
and morphisms $b_i : T'_i \to U'$ such
that $(g_i \circ a_{\alpha(i)}, b_i) : T'_i \to U \times_S U$
lifts to $R$.



\end{proof}





\begin{lemma}
\label{lemma-presentation-quasi-compact}
Let $S$ be a scheme. Let $U$ be a scheme over $S$.
Let $j = (s, t) : R \to U \times_S U$
be an etale equivalence relation on $U$ over $S$.
Then the quotient $U/R$
(see Groupoids, Definition \ref{groupoids-definition-quotient-sheaf})
is an algebraic space.
\end{lemma}

\begin{proof}
Let $U' \to U$ be a surjective, etale morphism.
Then $\{U' \to U\}$ is in particular an fppf covering.
Let $R'$ be the restriction of $R$ to $U'$, see
Groupoids, Definition \ref{groupoids-definition-restrict-relation}.
According to
Groupoids, Lemma \ref{groupoids-lemma-quotient-groupoid-restrict}
we see that $U/R \cong U'/R'$. Moreover, it is clear from the
description of $s', t'$ in
Groupoids, Lemma \ref{groupoids-lemma-restrict-groupoid}
that $s' , t' : R' \to U'$ are etale
as compositions of base changes of etale morphisms
(see Morphisms, Lemma \ref{morphisms-lemma-base-change-etale}
and \ref{morphisms-lemma-composition-etale}).
Thus we may replace $U$ by $U'$.

\medskip\noindent
We apply the previous remark to $U' = \coprod U_i$, where
$U = \bigcup U_i$ is an affine open covering with the property
that each $U_i$ maps into some affine open of $S$. Hence we
may and do assume that $U = \coprod U_i$ where
each $U_i$ maps into some affine open of $S$.

\medskip\noindent
Consider the restriction $R_i$ of $R$ to $U_i$.
It is clear that $\coprod F_i \to F$ is surjective.
We claim that each $U_i/R_i \to U/R$ is representable, and
an open immersion.







\end{proof}













\input{chapters}

\bibliography{my}
\bibliographystyle{alpha}

\end{document}
